Storm surges in the Strait of Georgia simulated with a regional model

### 2015-01-02 AO_guides/manuscriptAO.tex
Clarifified inversge barometer forcing

At the ocean surface, meteorological conditions and fresh water input from rivers are included. The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}. Forecasts from the period of 2002-2012 are available. Additionally, the inverse barometric effect of the atmospheric pressure is included, an important consideration in storm surge modelling. One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.

>>>>

At the ocean surface, meteorological conditions and fresh water input from rivers are included. The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}. Forecasts from the period of 2002-2012 are available. Forcing due to the local atmospheric pressure is also included by calculating an inverse barometer sea surface height,
\[ \eta_{IB} = -\frac{1}{g\rho_{0}}\left(P_{s}-P_0\right),\]
where $g$ is the accelaration due to gravity, $\rho_{0}$ is a reference density set to 1035 kg m$^{-3}$, $P_{s}$ is the sea level atmospheric pressure from the atmospheric model, and $P_0$ is a reference sea level pressure set to 101,000 Pa.  The gradient of $\eta_{IB}$  is scaled by $g$ and added to the forcing terms in the ocean momentum equations \citep{madec2012nemo}. As such, the contribution due to local atmospheric pressure is usually small since the spatial gradient in the sea level pressure is small compared to the model domain. Since the atmospheric model employs a terrain following vertical coordinate the model surface pressure has been corrected to sea level using \citep{holton1992introduction}
\[ P_s = P_1\left(\gamma\frac{z_1}{T_1} +1 \right)^\frac{g}{\gamma R},\]
where $P_1$ and $T_1$ are the model pressure and temperature at height $z_1$, $R$ is the ideal gas constant, and $\gamma$ is an temperature lapse rate set to 0.0098 K m$^{-1}$. One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.
